Capon Beamforming and Moving Objects - An Analysis of Lateral Shift-Invariance

### 2014-02-17 HRUIofMovingObjects/response_to_reviewers.tex

\documentclass{article}

\begin{document}

Answers to the reviewers in \textit{italic}.

\section*{Reviewer: 1}

As my scores indicate, I think the paper addresses an interesting problem and does it very well. Beside two VERY minor typos, I think it can be published unaltered.
1. "In" missing at the beginning of the Introduction section.
2. In section IIB, the opening sentence, there is an 's' missing in 'produce'.

\textit{We want to thank the reviewer for an encouraging review. Typos have been corrected.}

\section*{Reviewer: 2}

Thanks to the authors for addressing the problem of the temporal aspects in minimum variance beamforming, demonstrating the extra requirements on spatial sampling, and demonstrating how to efficiently oversample. The paper is good but I would like it to go through a second review round based on the following comments:
\\\\
Page2 column2 lines 34-35, and Fig. 1, L, M and d have not been introduced (need to have read the previous literature or equation (4) to know their meaning). I would suggest that in the introduction we do not need these details.
\\\\
Equation (8) the two factors p and q are confusing. Because q is always one, I would suggest omitting it.  Let us say that the oversampling is used both for Capon and for reducing folding of frequencies over the bandwidth.
\\\\
At what depth are the LSV plots taken with respect to the focal depth? Does it matter?
\\\\
Please specify the characteristics of the M5S probe (size, number of elements) if GE permits. Is it a phased or a linear array? What is the physical beam spacing, the field of view, the number of transmit and receive beams in the base (non oversampled) configuration? What is the transmit focal depth? Are transmit and receive apodization applied? Without this information, the work is not reproductible.
\\\\
In section “IV-A. Oversampling in transmit”: please confirm (specify) that 1x1 receive beamforming is performed at this point.
\\\\
In section “IV-B. Oversampling by parallel receive beamforming”, could you explain in more details what are the differences between Capon + STB and the technique proposed by Rabinovitch (2013) – reference [27]?
\\\\
Page10 Column1 line56  write “should be removed” instead of “should be remove”
\\\\
Page 10 Column 1 lines 40-60, you state that with synthetic transmit beamforming, some of the artifacts seen in oversampling in receive by phase rotation could be eliminated. It would be interesting to see such technique implemented.
\\\\
P10 column 2 line 28-29 write “too extreme” instead of “to extreme”
\\\\
Fig 6a: specify which beamforming is applied here (I suspect delay and sum with no multiline receive?). Maybe you want to consider showing still images of Capon and Capon-oversampled as well (not necessarily on fig. 6), just in case some people won’t look at the videos.

\end{document}

### 2014-02-17 HRUIofMovingObjects/jrnl_article.tex
Started answering comments from reviewers.

% Note that the a4paper option is mainly intended so that authors in
% countries using A4 can easily print to A4 and see how their papers will
% look in print - the typesetting of the document will not typically be
% affected with changes in paper size (but the bottom and side margins will).
% Use the testflow package mentioned above to verify correct handling of
% both paper sizes by the user's LaTeX system.
%
% Also note that the "draftcls" or "draftclsnofoot", not "draft", option
% should be used if it is desired that the figures are to be displayed in
% draft mode.
%
\documentclass[journal]{IEEEtran}
%
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
% \documentclass[journal]{../sty/IEEEtran}

@@

\newcommand\multimedia[1]{\textbf{{\color{red}#1}}}
\newcommand\comment[1]{\textit{{\color{red}(#1)}}}

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{Capon Beamforming Applied on Moving Objects - An Analysis of Local Lateral Shift-Invariance}
%
%
% author names and IEEE memberships
% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
% a structure at a ~ so this keeps an author's name from being broken across
% two lines.
% use \thanks{} to gain access to the first footnote area
% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
% was not built to handle multiple paragraphs
%

@@

In this paper we analyse the local lateral shift-invariance of the Capon beamformer when imaging moving objects. We then discuss and investigate different methods in order to improve on it. We show that insufficient lateral sampling makes an imaging system based on the Capon beamformer locally laterally shift-variant. We then show that the local lateral shift-invariance can be improved by oversampling based on phase rotation without affecting the acquisition frame rate and with minor changes in processing complexity.
\end{abstract}
% IEEEtran.cls defaults to using nonbold math in the Abstract.
% This preserves the distinction between vectors and scalars. However,
% if the journal you are submitting to favors bold math in the abstract,
% then you can use LaTeX's standard command \boldmath at the very start
% of the abstract to achieve this. Many IEEE journals frown on math
% in the abstract anyway.

@@

In \cite{Cox1973}, Cox presents in great detail how the amount of mismatch between the selected steering vector and the signal propagation vector impacts the beamformer output in simple situations. He also states the following important fact about the Capon beamformer [the $\vec{k}_3$ processor]: 
\begin{quote}
"\textit{... a $\vec{k}_3$ processor requires more closely spaced beams than a ... $\vec{k}_1$-processor in order to avoid serious signal suppression effects being introduced on signals arriving from directions between the beams}", 
\end{quote}
where $\vec{k}_1$ is the delay-and-sum (DAS) beamformer. In Fig.\,\ref{fig:das_capon_beams} we have plotted the two-way response for DAS and the Capon beamformer using (26) and (32) in \cite{Cox1973}. A 64-element linear array with $\lambda/2$ element spacing, where $\lambda$ is the wavelength, is scanned over a sector equal to four times the array resolution using narrow band excitation. Located in the farfield of the sector is an object moving parallel to the array surface. The plot shows the high precision scanned output when the object is located at plus and minus half the required beam spacing for DAS [See (\ref{eq:resolution})]. %No spatially white noise is present. Hence, the interference-noise matrix ($\sigma_0^2\mat{Q}$ in \cite{Cox1973}) is the identity. 
The effect of typical subarray averaging and diagonal loading ($L=32$ and $d=1/100$) have been added to the Capon beamformer output. We see that a DAS beamformer, using the required beam spacing, could yield a $1.9$ dB scalloping loss \cite{Harris1978} if the object is moving parallel to the array. However if the Capon beamformer is used with the same beam density the loss could be as large as $27$ dB. This imaging system is then highly sensitive to small spatial shifts, either by the object or the array itself. A $1.9$ dB scalloping loss for the Capon beamforming is obtained for this specific example when the beam spacing or object distance is decreased with a factor of 25.

@@

In previous literature on Capon beamforming for medical ultrasound imaging only single frame scenarios have been simulated. The scalloping loss effect has therefore been passed over in silence by either oversampling on transmit, by carefully positioning point scatterers exactly on transmit-receive beam pairs, or by ignorance. In our first work where Capon beamforming was applied on a loop of images containing moving objects \cite{Asen2012, Asen} we commented on the effect, but no solution was presented other than oversampling on transmit when simulating. The same was done by Jensen and Austeng \cite{Jensen2012}.

In this paper we investigate the issue further. We show that when a moving point scatterer is imaged with the Capon beamformer with the same beam spacing sufficient for DAS, the predicted scalloping loss is observed both in simulations and \textit{in-vitro} data. The question is then, what can we do in order to improve on the situation without affecting the acquisition frame rate and processing throughout. %We try two answer this by studying the scalloping loss of the Capon beamforming and investigating different methods aimed at reducing it. We also discuss how the different methods will affect the acquisition frame rate and the processing complexity.

The next section gives an introduction to Capon beamforming for medical ultrasound imaging together with a short introduction to lateral sampling and shift-invariance. In Section \ref{sec:capon_LLSI} we study the scalloping loss yield by the Capon beamformer and its local lateral shift-invariance property.  Different methods for reduced scalloping loss and by that improved local lateral shift-invariance are then presented in Section \ref{sec:methods}. Then, in Section \ref{sec:res} we present the results of applying the most promising method on \textit{in-vitro} data. Finally we discuss our findings in Section \ref{sec:dis} and give our conclusions in Section \ref{sec:con}.

@@

\subsection{Capon Beamforming}
The Capon beamformer produce a weight vector, like the one in (\ref{eq:das}), based on the impinging signal and an optimization criterion. The problem is as follow \cite{Capon1969}:
\begin{align}
&\min_{\vec{w}} E\{|z[n]|^2\} \rightarrow \min_{\vec{w}} \vec{w}^H \mat{\hat{R}} \vec{w} \label{eq:capon_optimization_criteria} \\
&\text{subjected to } \vec{w}^H\vec{a} = 1,
\end{align}
where $\mat{\hat{R}}$ is a sample covariance matrix. Hence, the resulting weight vector will minimizes the output power while maintaining unit gain in the steering direction $\vec{a}$. %The covariance matrix is usually unknown and has to be estimated from the element data. The covariance matrix in (\ref{eq:capon_optimization_criteria}) is therefore substituted with the sample covariance matrix.

@@

The lateral sampling required in an ultrasound image for a given center frequency, $f_c$, and bandwidth, $B$, is given by the aperture size, $D$, and the wavelength of the center frequency, $\lambda_c = c/f_c$, where $c$ is the speed of sound. The required Nyquist beam spacing is then \cite{Hergum2007}:
\begin{align}
\Delta < \frac{\lambda_c}{qpD}, \label{eq:resolution}
\end{align}
where the oversampling factor $p$ is applied to reduce folding of frequencies between $f_c$ and $f_c + B/2$. A $p$-factor of two has been used throughout this paper. The oversampling factor $q$ will be used to oversample prior to Capon beamforming and is one if not stated otherwise.

@@

In Fig.\,\ref{fig:das_shiftvariant} the beam density is reduced by a factor of two compared to Fig.\,\ref{fig:das_shiftinvariant}. The geometric distortion is now increased to a level where the contours are not diagonal. The maximum scalloping loss is also larger than 1 dB. Hence, the imaging system is no longer locally laterally shift-invariant. We observe that the maximum peak is both geometrically distorted and attenuated when the point is located between two beams and the distance between the beams is too high.

\begin{figure*}[!t]
\centerline{
\subfloat[]{\includegraphics[width=0.5\linewidth]{\img fig2a.eps}\label{fig:das_shiftinvariant}}
\hfill{}
\subfloat[]{\includegraphics[width=0.5\linewidth]{\img fig2b.eps}\label{fig:das_shiftvariant}}
}
\caption{LSV-plots of delay-and-sum beamforming. If the contours are diagonal and the amplitude variation among the peak points (black dots) is small, the imaging system is said to be locally laterally shift-invariant. Dotted vertical lines are plotted where transmit-receive beam pairs are located. a) Sampling density according to (\ref{eq:resolution}). The system is locally laterally shift-invariant. b) Undersampling with a factor of two. The system is no longer locally laterally shift-invariant.}
\label{fig:das}
\end{figure*}

@@

The diagonal factor $d$ and the effective adaptive aperture size $L$ can be used to control the mainlobe width of the Capon beamformer. In both extreme cases, $d\gg1$ and $L=1$, the Capon beamformer becomes either a triangle or uniform-weighted DAS beamformer respectively, and a beam spacing according to (\ref{eq:resolution}) would then be sufficient. In Fig.\,\ref{fig:capon2} we show how the scalloping loss is reduced when the amount of diagonal loading is increased.

We observe that Fig.\,\ref{fig:capon2} now contains a scaled version of the pattern presented in Fig.\,\ref{fig:das_shiftvariant}. From Fig.\,\ref{fig:capon} it should be obvious that the Capon beamformer either requires increased lateral sampling in order to make the imaging system locally laterally shift-invariant, or the resolution has to be reduced to the same level as DAS. It would be unwise to do all the computation involved with Capon beamforming with a high $d$ value and end up with a resolution equal to DAS. However, we see that the diagonal factor can be a useful tool if resolution has to be traded for the need of oversampling. The same is true for the subarray length parameter.

\begin{figure*}[!t]
\centerline{
\subfloat[]{
\includegraphics[width=0.5\linewidth]{\img fig3a.eps}\label{fig:capon1}
}
\subfloat[]{
\includegraphics[width=0.5\linewidth]{\img fig3b.eps}\label{fig:capon2}
}
}
\caption{LSV-plots of Capon beamforming with $L=M/2=48$ and $K=1$, and with sampling corresponding to (\ref{eq:resolution}). a) Capon beamforming with $d=1/100$ as diagonal loading. The system is not locally lateral shift-invariant. b) The diagonal loading factor $d$ is increased to 1. The scalloping loss is reduced, but the system is still not locally lateral shift-invariant.}
\label{fig:capon}
\end{figure*}

\section{Oversampling Methods}\label{sec:methods}
In this section we will discuss different methods aimed at improving the local lateral shift-invariance for the Capon beamformer. From the previous section it should be clear that we somehow need to increase the lateral sampling to make the Capon imaging system locally laterally shift-invariant. In this section we will investigate how this oversampling should be conducted, how large it needs to be and how the proposed methods will affect the acquisition frame rate and the processing complexity.

\subsection{Oversampling on Transmit}
A straight forward way of obtaining sufficient sampling in an active imaging system is by increasing the number of transmit beams (tx-beams) until the system is locally laterally shift-invariant. In Fig.\,\ref{fig:capon_oversampling_a} and Fig.\,\ref{fig:capon_oversampling_b} we have increased the lateral sampling on transmit by a $q$-factor of four and sixteen respectively. As expected the geometric distortion and scalloping loss decreases with increased lateral sampling. Note that Fig.\,\ref{fig:capon_oversampling_b} is zoomed compared to Fig.\,\ref{fig:capon} in order to better see the beam-to-beam variation.

From Fig.\,\ref{fig:capon_oversampling_b} we see that sixteen times oversampling reduces the scalloping loss significantly, but the maximum is still larger than 1 dB. A even higher level of oversampling is therefore required in order to have a completely local lateral shift-invariant Capon beamformer when a moving point scatterer is simulated without noise and with the same parameters as in Fig\,\ref{fig:capon}. %$1/100$ in diagonal loading. 
Note that a diagonal loading factor of $1/100$ will introduce a white noise component in the sample covariance matrix $20$ dB lower than the signal present. Hence, the covariance has been estimated from data which appeared to have an SNR of 20 dB.

@@

\begin{figure*}[!t]
\centerline{
\subfloat[]{
\includegraphics[width=0.5\linewidth]{\img fig4a.eps}\label{fig:capon_oversampling_a}
}
\subfloat[]{
\includegraphics[width=0.5\linewidth]{\img fig4b.eps}\label{fig:capon_oversampling_b}
}
}
\caption{Capon beamforming as described in Fig\,\ref{fig:capon} with oversampling on transmit. The figure is zoomed compared with Fig\,\ref{fig:capon} to better see the beam-to-beam variation. a) Four times oversampling. b) Sixteen times oversampling. Even though its reduced, the maximum scalloping loss is still too high for the system to be named locally laterally shift invariant.}
\label{fig:capon_oversampling}
\end{figure*}

\subsection{Oversampling by Parallel Receive Beamforming}
Parallel receive beamforming (PRB) means that multiple narrow receive beams (rx-beams) are constructed from one broad tx-beam. Then, by reducing the number of tx-beams and relying on multiple rx-beams in order to get sufficient sampling on receive, it is possible to achieve high frame rates in e.g. 3D cardiac ultrasound imaging. However, PRB also introduces geometrical distortion which to some degree can be compensated for by using e.g. synthetic transmit beamforming (STB) \cite{Hergum2007, Denarie2011, Denarie2013}. In \cite{Rabinovich2013} it is shown that STB works well with Capon beamforming. In \cite{Asen} we also used STB in order to increase the beam density of a cardiac ultrasound recording. In the following discussion we consider corrected PRB as being equal to oversampling on transmit when it comes to image quality. The LSV-plots for PRB are therefore not included.

In cardiac ultrasound imaging, PRB is typically used in order to keep the number of rx-beams high while the number of tx-beams are reduced. When using PRB for oversampling we would like to maintain the same number of tx-beams while the number of rx-beams are increased. The acquisition frame rate will then not be affected. On the other hand more rx-beams will increase the number of covariance matrices, and thereby the number of matrix-inversions with a factor equal to the number of parallel receive lines. The processing times reported in \cite{Asen} will then also increase with the same factor. Since more than sixteen times oversampling is required in order to achieve lateral-shift-invariant and high-resolution imaging of a moving point scatterer, none of the implementations described in \cite{Asen} would be real time any more. We therefore need a different way of oversampling our imaging sector.

\subsection{Oversampling by Phase Rotation}
The Capon beamformer formula in (\ref{eq:capon_weights}) gives a direct opportunity for oversampling using phase rotation. The steering vector $\vec{a}$ has in previous work been set to $\vec{1}$ since ultrasound data have to be pre-delayed. However, the steering vector can also, as in narrow band applications, be varied over a set of pre-defined vectors. In a narrow band setting the beam can make a sweep across the whole imaging sector by phase steering. This is also possible in broadband applications as long as the phase rotation is less than one pulse length (See coarse-fine beamforming in \cite{Thomenius}). The maximal steering angle is therefore given by
\begin{align}
\theta_{max} = \alpha p \Delta,
\end{align}
where $\alpha$ is the pulse length in $\lambda$'s, and $p$ and $\Delta$ is the oversampling factor and beam spacing respectively as defined in (\ref{eq:resolution}). For our simulations $\theta_{max} = 3\Delta$.
It is therefore possible to make a sweep using phase rotation between a given beam and half-way to its two nearest neighbours without introducing to large errors. Phase rotation is actually what the Capon beamformer use internally to micro-steer away from bright point scatterers and to enhance edges.

@@

@@

\begin{figure*}[!t]
	\centerline{
		\subfloat[]{
			\includegraphics[width=0.5\linewidth]{\img fig5a.eps}
		}
		\subfloat[]{
			\includegraphics[width=0.5\linewidth]{\img fig5b.eps}
		}
	}
	\caption{Capon beamforming as described in Fig\,\ref{fig:capon} with oversampling based on phase rotation. The figure is zoomed compared with Fig\,\ref{fig:capon} to better see the beam-to-beam variation. a) Four times oversampling. b) Sixteen times oversampling.}
	\label{fig:benchmark_capon_bs}
\end{figure*}

@@

In Fig\,\ref{fig:image_das} one point scatterer has been selected (white box), and then used to generate LSV-plots. Figure \ref{fig:phantom_das} shows the resulting LSV-plot when using the DAS beamformer. The scalloping loss is less than 1 dB, and the system is therefore locally laterally shift-invariant. As a consequence point scatterers have a consistent appearance when subjected to motion \multimedia{Video DAS}.

Figure \ref{fig:phantom_capon_LSV} shows the same point scatterer response processed with the Capon beamformer without oversampling. For the frame where the point scatterer is at -6 degrees the receive beam direction is very close to the signal propagation vector. Hence, we see a high amplitude value at this location. In almost all the other frames the signal loss is massive (at most 10 dB). The result is blinking point scatterers when subjected to motion \multimedia{Video Capon}.

>>>>

% Note that the a4paper option is mainly intended so that authors in
% countries using A4 can easily print to A4 and see how their papers will
% look in print - the typesetting of the document will not typically be
% affected with changes in paper size (but the bottom and side margins will).
% Use the testflow package mentioned above to verify correct handling of
% both paper sizes by the user's LaTeX system.
%
% Also note that the "draftcls" or "draftclsnofoot", not "draft", option
% should be used if it is desired that the figures are to be displayed in
% draft mode.
%
\documentclass[draftcls]{IEEEtran}
%\documentclass[journal]{IEEEtran}
%
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
% \documentclass[journal]{../sty/IEEEtran}

@@

\newcommand\multimedia[1]{\textbf{{\color{red}[#1]}}}
\newcommand\comment[1]{\textit{{\color{red}(#1)}}}

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{Capon Beamforming and Moving Objects - An Analysis of Lateral Shift-Invariance}
%
%
% author names and IEEE memberships
% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
% a structure at a ~ so this keeps an author's name from being broken across
% two lines.
% use \thanks{} to gain access to the first footnote area
% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
% was not built to handle multiple paragraphs
%

@@

In this paper we analyse the local lateral shift-invariance of the Capon beamformer when imaging moving objects. We show that insufficient lateral sampling makes an imaging system based on the Capon beamformer laterally shift-variant. Different methods for oversampling on transmit and receive are then discussed and investigated in order to improve on it. It is shown that lateral shift-invariance can be improved by oversampling based on phase rotation on receive. This without affecting the acquisition frame rate and with a minor change in processing complexity.
\end{abstract}
% IEEEtran.cls defaults to using nonbold math in the Abstract.
% This preserves the distinction between vectors and scalars. However,
% if the journal you are submitting to favors bold math in the abstract,
% then you can use LaTeX's standard command \boldmath at the very start
% of the abstract to achieve this. Many IEEE journals frown on math
% in the abstract anyway.

@@

In \cite{Cox1973}, Cox presents in great detail how the amount of mismatch between the selected steering vector and the signal propagation vector impacts the beamformer output in simple situations. He also states the following important fact about the Capon beamformer [the $\vec{k}_3$ processor]: 
\begin{quote}
"\textit{... a $\vec{k}_3$ processor requires more closely spaced beams than a ... $\vec{k}_1$-processor in order to avoid serious signal suppression effects being introduced on signals arriving from directions between the beams}", 
\end{quote}
where $\vec{k}_1$ is the delay-and-sum (DAS) beamformer. In Fig.\,\ref{fig:das_capon_beams} we have plotted the two-way response for DAS and the Capon beamformer using (26) and (32) in \cite{Cox1973}. A 64-element linear array with $\lambda/2$ element spacing, where $\lambda$ is the wavelength, is scanned over a sector equal to four times the array resolution using narrow band excitation. Located in the farfield of the sector is an object moving parallel to the array surface. The plot shows the high precision scanned output when the object is located at plus and minus half the required beam spacing for DAS [See (\ref{eq:resolution})]. %No spatially white noise is present. Hence, the interference-noise matrix ($\sigma_0^2\mat{Q}$ in \cite{Cox1973}) is the identity. 
The effect of typical subarray averaging and diagonal loading ($L=32$ and $d=1/100$) have been added to the Capon beamformer output. We see that a DAS beamformer, using the required beam spacing, could yield a $1.9$ dB scalloping loss \cite{Harris1978} if the object is moving parallel to the array. However, if the Capon beamformer is used with the same beam density the loss could be as large as $27$ dB. This imaging system is then highly sensitive to small spatial shifts, either by the object or the array itself. A $1.9$ dB scalloping loss for the Capon beamforming is obtained for this specific example when the beam spacing or object distance is decreased with a factor of 25.

@@

In previous literature on Capon beamforming for medical ultrasound imaging only single frame scenarios have been simulated. The scalloping loss effect has therefore been passed over in silence by either oversampling on transmit, by carefully positioning point scatterers exactly on transmit-receive beams, or by ignorance. In our first work where Capon beamforming was applied on a loop of images containing moving objects \cite{Asen2012, Asen} we commented on the effect, but no solution was presented other than oversampling on transmit when simulating. The same was done by Jensen and Austeng \cite{Jensen2012}.

In this paper we investigate the issue further. We show that when a moving point scatterer is imaged with the Capon beamformer with the same beam spacing which is sufficient for DAS, the predicted scalloping loss is observed both in simulations and \textit{in-vitro} data. The question is then, what can we do in order to improve on the situation without affecting the acquisition frame rate and processing throughout. %We try two answer this by studying the scalloping loss of the Capon beamforming and investigating different methods aimed at reducing it. We also discuss how the different methods will affect the acquisition frame rate and the processing complexity.

The next section gives an introduction to Capon beamforming together with a short introduction to lateral sampling and shift-invariance. In Section \ref{sec:capon_LLSI} we study the scalloping loss yield by the Capon beamformer and its local lateral shift-invariance property.  Different methods for reduced scalloping loss and by that improved local lateral shift-invariance are presented in Section \ref{sec:methods}. Then, in Section \ref{sec:res} we present the results of applying the most promising method on \textit{in-vitro} data. Finally we discuss our findings in Section \ref{sec:dis} and give our conclusions in Section \ref{sec:con}.

@@

\subsection{Capon Beamforming}
The Capon beamformer produces a weight vector, like the one in (\ref{eq:das}), based on the impinging signal and an optimization criterion. The problem is as follow \cite{Capon1969}:
\begin{align}
&\min_{\vec{w}} E\{|z[n]|^2\} \rightarrow \min_{\vec{w}} \vec{w}^H \mat{\hat{R}} \vec{w} \label{eq:capon_optimization_criteria} \\
&\text{subjected to } \vec{w}^H\vec{a} = 1,
\end{align}
where $\mat{\hat{R}}$ is a sample covariance matrix. Hence, the resulting weight vector minimizes the output power while maintaining unit gain in the steering direction $\vec{a}$. %The covariance matrix is usually unknown and has to be estimated from the element data. The covariance matrix in (\ref{eq:capon_optimization_criteria}) is therefore substituted with the sample covariance matrix.

@@

The required lateral sampling in an ultrasound image for a given center frequency, $f_c$, and bandwidth, $B$, is given by the aperture size, $D$, and the wavelength of the center frequency, $\lambda_c = c/f_c$, where $c$ is the speed of sound. The required Nyquist beam spacing is then \cite{Hergum2007}:
\begin{align}
\Delta < \frac{\lambda_c}{qpD}, \label{eq:resolution}
\end{align}
where the oversampling factor $p$ is applied to reduce folding of frequencies between $f_c$ and $f_c + B/2$. A $p$-factor of two has been used throughout this paper. The oversampling factor $q$ will be used to oversample prior to Capon beamforming and is one if not stated otherwise.

@@

In Fig.\,\ref{fig:das_shiftvariant} the beam density is reduced by a factor of two compared to Fig.\,\ref{fig:das_shiftinvariant}. The geometric distortion is now increased to a level where the contours are not diagonal. The maximum scalloping loss is almost 4 dB. Hence, the imaging system is no longer locally laterally shift-invariant. We observe that the maximum peak is both geometrically distorted and attenuated when the point is located between two beams and the distance between the beams is too high.

\begin{figure*}[!t]
\centerline{
\subfloat[]{\includegraphics[width=0.5\linewidth]{\img fig2a.eps}\label{fig:das_shiftinvariant}}
\hfill{}
\subfloat[]{\includegraphics[width=0.5\linewidth]{\img fig2b.eps}\label{fig:das_shiftvariant}}
}
\caption{LSV-plots of delay-and-sum beamforming. If contours are diagonal and the amplitude variation among the peak points (black dots) is small, the imaging system is said to be locally laterally shift-invariant. Dotted vertical lines are plotted where transmit-receive beams are located. a) Sampling density according to (\ref{eq:resolution}). The system is locally laterally shift-invariant. b) Undersampling by a factor of two. The system is no longer locally laterally shift-invariant.}
\label{fig:das}
\end{figure*}

@@

The diagonal factor $d$ and the effective adaptive aperture size $L$ can be used to control the mainlobe width of the Capon beamformer. In the extreme cases, $d\gg1$ and $L=1$, the Capon beamformer becomes either a triangle or uniform-weighted DAS beamformer respectively, and a beam spacing according to (\ref{eq:resolution}) would then be sufficient. In Fig.\,\ref{fig:capon2} we show how the scalloping loss is reduced when the amount of diagonal loading is increased.

We observe that Fig.\,\ref{fig:capon2} now contains a scaled version of the pattern presented in Fig.\,\ref{fig:das_shiftvariant}. From Fig.\,\ref{fig:capon} it should be clear that Capon beamforming either requires increased lateral sampling, or the resolution has to be reduced to the same level as DAS in order to make the imaging system locally laterally shift-invariant. It would be unwise to do all the computation involved with Capon beamforming with a high $d$ value and end up with a resolution equal to DAS. However, we see that diagonal loading can be a useful tool if resolution has to be traded for the need of oversampling. The same is true for the subarray length parameter.

\begin{figure*}[!t]
\centerline{
\subfloat[]{
\includegraphics[width=0.5\linewidth]{\img fig3a.eps}\label{fig:capon1}
}
\subfloat[]{
\includegraphics[width=0.5\linewidth]{\img fig3b.eps}\label{fig:capon2}
}
}
\caption{LSV-plots of Capon beamforming with $L=M/2=48$ and $K=1$, and with sampling corresponding to (\ref{eq:resolution}). a) Capon beamforming with $d=1/100$ as diagonal loading. The system is not locally laterally shift-invariant. b) The diagonal loading factor $d$ is increased to 1. The scalloping loss is reduced, but the system is still not locally laterally shift-invariant.}
\label{fig:capon}
\end{figure*}

\section{Oversampling Methods}\label{sec:methods}
In this section we will discuss different methods aimed at improving the local lateral shift-invariance of the Capon beamformer. From the previous section it should be clear that we somehow need to increase the lateral sampling to make the Capon imaging system locally laterally shift-invariant. In this section we will investigate how this oversampling should be conducted, how large it needs to be and how the proposed methods will affect the acquisition frame rate and the processing complexity.

\subsection{Oversampling on Transmit}
A straight forward way of obtaining sufficient sampling in an active imaging system is by increasing the number of transmit beams (tx-beams) until the system is locally laterally shift-invariant. In Fig.\,\ref{fig:capon_oversampling_a} and Fig.\,\ref{fig:capon_oversampling_b} we have increased the tx-beam density by a $q$-factor of four and sixteen respectively. As expected the geometric distortion and scalloping loss decreases with increased lateral sampling. Note that Fig.\,\ref{fig:capon_oversampling_b} is zoomed compared to Fig.\,\ref{fig:capon} in order to better see the beam-to-beam variation.

From Fig.\,\ref{fig:capon_oversampling_b} we see that sixteen times oversampling reduces the scalloping loss significantly. Even though high frequency variations are reduced, the maximum is still larger than 1 dB. An even higher level of oversampling is therefore required in order to have a completely local lateral shift-invariant Capon beamformer when a moving point scatterer is simulated without noise and with the same parameters as in Fig\,\ref{fig:capon}. %$1/100$ in diagonal loading. 
Note that a diagonal loading factor of $1/100$ will introduce a white noise component in the sample covariance matrix $20$ dB lower than the signal present. Hence, the covariance has been estimated from data which appeared to have an SNR of 20 dB.

@@

\begin{figure*}[!t]
\centerline{
\subfloat[]{
\includegraphics[width=0.5\linewidth]{\img fig4a.eps}\label{fig:capon_oversampling_a}
}
\subfloat[]{
\includegraphics[width=0.5\linewidth]{\img fig4b.eps}\label{fig:capon_oversampling_b}
}
}
\caption{Capon beamforming as described in Fig\,\ref{fig:capon} with oversampling on transmit. The figure is zoomed compared with Fig\,\ref{fig:capon} to better see the beam-to-beam variation. a) Four times oversampling. b) Sixteen times oversampling. The maximum scalloping loss is reduced, but it is still too high for the system to be locally laterally shift invariant.}
\label{fig:capon_oversampling}
\end{figure*}

\subsection{Oversampling by Parallel Receive Beamforming}
Parallel receive beamforming (PRB) means that multiple narrow receive beams (rx-beams) are constructed from one broad tx-beam. Then, by reducing the number of tx-beams and relying on multiple rx-beams in order to get sufficient sampling on receive, it is possible to achieve high frame rate imaging. However, PRB also introduces geometrical distortion which to some degree can be compensated for by using e.g. synthetic transmit beamforming (STB) \cite{Hergum2007, Denarie2013}. In \cite{Rabinovich2013} it is shown that STB works well with Capon beamforming. In \cite{Asen} we also used STB in order to increase the beam density of a cardiac ultrasound recording. In the following discussion we consider corrected PRB as being equal to oversampling on transmit when it comes to image quality. The LSV-plots for PRB are therefore not included.

In cardiac ultrasound imaging, PRB is typically used in order to keep the number of rx-beams high while the number of tx-beams are reduced. When using PRB for oversampling we would like to maintain the same number of tx-beams while the number of rx-beams are increased. The acquisition frame rate will then not be affected. On the other hand, more rx-beams will increase the number of covariance matrices and thereby the number of matrix-inversions with a factor equal to the number of parallel receive lines. The processing times reported in \cite{Asen} will then also increase with the same factor. Since more than sixteen times oversampling is required in order to achieve lateral-shift-invariant and high-resolution imaging of a moving point scatterer, none of the implementations described in \cite{Asen} would be real time any more. We therefore need a different way of oversampling our imaging sector.

\subsection{Oversampling by Phase Rotation}
The Capon beamformer formula in (\ref{eq:capon_weights}) gives a direct opportunity for oversampling using phase rotation. The steering vector $\vec{a}$ has in previous work been set to $\vec{1}$ since ultrasound data have to be pre-delayed. However, the steering vector can also, as in narrow band applications, be varied over a set of pre-defined vectors. In a narrow band setting the beam can make a sweep across the whole imaging sector through phase steering. This is also possible in broadband applications as long as the phase rotation is less than one pulse length (See coarse-fine beamforming in \cite{Thomenius} and its references). The maximal steering angle is approximately given by
\begin{align}
\theta_{max} = \alpha p \Delta,
\end{align}
where $\alpha$ is the pulse length in $\lambda$'s, and $p$ and $\Delta$ is the oversampling factor and beam spacing respectively as defined in (\ref{eq:resolution}). For our simulations $\theta_{max} = 3\Delta$.
It is therefore possible to make a sweep based on phase rotation between a given beam and half-way to its two nearest neighbours without introducing large errors. Phase rotation is actually what the Capon beamformer uses internally to micro-steer away from bright point scatterers and to enhance edges.

@@

Figure \ref{fig:capon_oversampling_pr} shows the phase rotation method applied on the same scan sequence as in Fig\,\ref{fig:capon} with four and sixteen times oversampling respectively. Note that the figure is zoomed compared with Fig\,\ref{fig:capon}. As with oversampling on transmit there is some variation when the point moves in and out of focus. The maximum scalloping loss is therefore large. The extreme local variations are however reduced significantly. Comparing Fig\,\ref{fig:capon_oversampling} with Fig\,\ref{fig:capon_oversampling_pr}, the errors introduced by the phase rotation method are best seen in sidelobes.

@@

\begin{figure*}[!t]
	\centerline{
		\subfloat[]{
			\includegraphics[width=0.5\linewidth]{\img fig5a.eps}
		}
		\subfloat[]{
			\includegraphics[width=0.5\linewidth]{\img fig5b.eps}
		}
	}
	\caption{Capon beamforming as described in Fig\,\ref{fig:capon} with oversampling based on phase rotation. The figure is zoomed compared with Fig\,\ref{fig:capon} to better see the beam-to-beam variation. a) Four times oversampling. b) Sixteen times oversampling.}
	\label{fig:capon_oversampling_pr}
\end{figure*}

@@

One point scatterer in Fig\,\ref{fig:image_das} has been selected [white box], and then used to generate LSV-plots. Figure \ref{fig:phantom_das} shows the resulting LSV-plot when using the DAS beamformer. The scalloping loss is less than 1 dB, and the system is by that locally laterally shift-invariant. As a consequence point scatterers have a consistent appearance when subjected to motion \multimedia{Media Movie 1}.

Figure \ref{fig:phantom_capon_LSV} shows the same point scatterer response processed with the Capon beamformer without oversampling. For the frame where the point scatterer is at -6 degrees the receive beam direction is very close to the signal propagation vector. Hence, we see a high amplitude value at this location. In almost all the other frames the signal loss is massive (at most 10 dB). The result is blinking point scatterers when subjected to motion \multimedia{Media Movie 2}.
